How to Recognize a Bad ``Bad Smell'' using Contrast Sets

### 2016-03-09 Manuscript.tex
Updates from ShareLaTeX

\title{  How to Recognize  a Bad  ``Bad Smell'' (using Contrast Sets)}
\numberofauthors{2} 
\author{  
\alignauthor
Rahul~Krishna, Tim Menzies  \\
       \affaddr{Computer Science, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies\}@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE, College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract} 
It is hard to assess
if a bad smell is ``bad''; i.e. if it is irrelevant or misleading
for a specific project.
Developers, text books, and tools
disagree on which bad smells are important.
Specifically, it is unclear at what threshold values
for which metrics should trigger code reorganization since
most studies do not offer
metrics {\em and} a threshold for their proposed bad smells.
This paper proposes the XTREE contrast set
learner to assess bad smells
and reject any that are not supported by
evidence in the historical record.  
Using XTREE, we   show that the range of useful
``bad smells'' changes from project to project. Hence,
before refactoring, it is important to run a critique tool
like XTREE to stop work on useless code reorganizations.
Further, using XTREE, it is possible to show that many
of the bad smells endorsed by standard texts or tools
are not useful for particular projects. 
Based on these results, we endorse  
guiding code reorganization via bad smells. However,  
it is important to XTREE the bad smells defined for a 
particular project.

@@

This paper proposes the XTREE contrast set learner for recognizing
``bad'' bad smells. 
Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code reorganization. Hence, constrast set learners
like XTREE can be used to  recognize ``bad''
bad smells:
\begin{itemize}
    \item Consider $N$ developers each proposing a software reorganization
based on a set of bad smells $S$.
\item XTREE will (a)~learn contrast sets
from the historical record; then (b)~reject  any reorganization based on ``bad'' bad smells, i.e.
 on any member of $S_{\mathit{bad}} \in S$
that XTREE could {\em not} be found in the historical record.
\end{itemize}
Note that we    recommend  XTREE  is used to critize, but  not to create,
bad smells. 
XTREE uses frequency counts of code features seen in the historical record
to generate
predictions about the effects of changing code
attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
where spurious correlations can suggest effects that are not
truly causal.

@@

Nevertheless, XTREE is a still a useful tool to critique and prune
irrelevant bad smells since, 
as show below, many of the bad smell detectors offered
in the literature (and supported by readily available tools)
fall outside of  space of
effects found  using XTREE's analysis.
That is,   XTREE can find 
{\em neither} correlation {\em nor}
causal evidence that those bad smells
have any support   in the historical record.

@@

\newpage
The contributions of this paper are
\begin{itemize}

@@

@@

To support the last paragraph, we offer several
specific studies
showing that   widely-held views in SE that are   now questionable, given recent results. 
The first few come from the general SE literature while the rest
relate specifically to bad smells:
\input{tex/refactor}
\begin{enumerate} 
\item
Nagappan et al. analyzed~\cite{mei15} 
how   {\tt goto}  was used  in 11,000 Github repositories. 
They found that developers
rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
\item
In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em fixing a requirements error  during
coding can be thousands of times more expensive than recognizing and fixing that error
as soon as it arises}. 
We found no evidence for such large   phase delay effect in modern
software (in 171 software projects shepherded
by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ software development languages, tools and development practices. 
Whatever the reason, the main point  is that phase delay is a widely
held belief, despite little recent evidence to support it.
\item
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, They found that  ``(a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
\end{enumerate}
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
careers.

@@

\section{Learning Bad Smell Thresholds}

After a review of the literature, we reached the same conclusion
as Shatnawi~\cite{Shatnawi10}; i.e. there is a dearth of papers recommending specific thresholds
for bad smells. In the recent literature (since 2010), we could
only find the handful   papers discussed below.

In the recent literature, there are two approaches for
learning bad smell thresholds:
\begin{itemize}
    \item Approaches based on {\em outlier statistics}~\cite{erni96,bender99}
    as used by Shatnawi~\cite{shatnawi10}, Alvenes et al.~\cite{Alves2010}
    and Hermans et al.~\cite{hermans15};
    \item  Approaches based on {\em cluster deltas} that we developed
    for   Centroid Deltas~\cite{me12c} and XTREE.
\end{itemize}
The outlier approach assume that unusually large measurements are risk-prone.
Hence, they generate one bad smell threshold for any metric
with such an unusually large measurements.

The advantage of outlier
methods is that they are simple to implement. However,
their disadvantage are two-fold. 
Firstly, the outlier approach is {\em verbose}; i.e.  it can
propose thresholds on most of the code metrics.
Such verbosity makes it hard   to critique and prune away proposed code reorganizations since any bad smell mentioned by a developer has some support
somewhere amongst all those proposed thresholds.

Secondly, they do not reflect
on sets of code metrics that change together. For example,
one way to remove unusually large methods is to reorganize that method's
functionality into several smaller ones. If that is done,
this means that {\em reducing} lines of code in one method
will {\em increase} the coupling of the methods of that class to
either itself or to other classes (depending on whether or not
the new smaller methods are placed in the original class, or in other classes).
Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
large lines of code {\em and} seemingly excessive coupling.

Borges and Menzies first proposed the centroid delta approach to address
this second problem. Centroid deltas assume...

One is {\em succinct}; i.e. comments
on too many code metrics. If the

class' coupling mea
\begin{itemize}
    \item 
\end{itemize}that they do not consider the associations
between variables.

Cluster deltas methods assume that at
Prior work on deriving metric thresholds from benchmark data
used {\em outlier} statistics to identify abnormally large ranges. 
Another approach, used by

For example,
one of the first methods of finding thresholds was proposed Erni and Lewerentz~\cite{erni96}. Their technique to identify thresholds was based on the data distribution, specifically the mean and the standard deviation of the metric values. They propose the use of values that lie beyond one standard deviation from the mean as a threshold. The minimum value $T_{min}$ is given by $T_{min}=\mu-\sigma$, and this is used when metric definition considers very small values as an indicator of problems. Otherwise, $T_{max}=\mu+\sigma$ is used, when large metric values are considered problematic.

Several researchers~\cite{shatnawi10}~cite{alves10} have pointed out that this method is subject to a few problems. Firstly, it doesn't consider the fault-proneness of classes when the thresholds are computed. Secondly, there is a lack of empirical validation of this methodology, which impedes reasonable comparisons.

@@

>>>>

\title{  How to Recognize  a Bad  ``Bad Smell'' using Contrast Sets}
\numberofauthors{2} 
\author{  
\alignauthor
Rahul~Krishna, Tim Menzies  \\
       \affaddr{Computer Science, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies\}@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE, College Park, USA}\\ 
       {llayman@cese.fraunhofer.org}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract} 
It is hard to assess
if a ``bad smell'' is irrelevant or misleading
for a specific project.
Developers, text books, and tools
disagree on which bad smells are important. 
Specifically, it is unclear at what threshold values
for which metrics should trigger code reorganization\footnote{From LL: The term reorganization is a bit weird. Bad code smells are ``refactored''. Do you have a reason that you are using reorganization instead of refactoring throughout the paper? Recall that Fowler's code smell book is called ``Refactoring''} since
most studies do not offer
both metrics {\em and} a threshold for their proposed bad smells.
This paper proposes the XTREE contrast set
learner to assess bad smells
and reject any that are not supported by
evidence in the historical record.  
Using XTREE, we   show that the range of useful
``bad smells'' changes from project to project. Hence,
before refactoring, it is important to run a critique tool
like XTREE to stop work on useless code reorganizations.
Further, XTREE shows that many
of the bad smells defined in standard texts and tools
are not useful for particular projects. 
Based on these results, we endorse  
guiding code reorganization via bad smells, but one must empirically validate the importance of specific bad smells within the context of each project using an approach like XTREE.

@@

This paper proposes the XTREE contrast set learner for recognizing
``bad'' bad smells. 
Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code reorganization. Hence, constrast set learners
like XTREE can be used to  recognize ``bad''
bad smells:
\begin{itemize}
    \item Consider $N$ developers each proposing a software reorganization 
based on a set of bad smells $S$.
\item XTREE will (a)~learn contrast sets
from the historical record; then (b)~reject  any reorganization based on ``bad'' bad smells, i.e.
 on any member of $S_{\mathit{bad}} \in S$
that XTREE could {\em not} be found in the historical record.
\end{itemize}
Note that we    recommend  XTREE  is used to critize, but  not to create,
bad smells. 
XTREE uses frequency counts of code features seen in the historical record\footnote{From LL: Note that ``new'' bad smells may not have a historical record. A limitation to be acknowledged and a reason why developer knowledge is still relevant.}
to generate
predictions about the effects of changing code
attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
where spurious correlations can suggest effects that are not
truly causal.

@@

Nevertheless, XTREE is a still a useful tool to critique and prune
irrelevant bad smells since, 
as show below, many of the bad smell detectors offered
in the literature (and supported by readily available tools)
fall outside of  space of
effects found  using XTREE's analysis.
That is,   XTREE can find 
{\em neither} correlation {\em nor}
causal evidence that those bad smells
have any support   in the historical record.\footnote{From LL: You need to say what ``support from the historical record'' is early in the paper. Are you mapping code smells in source files to defects? Code smells to performance regressions? Whatever it is, give an example.}

@@

The contributions of this paper are
\begin{itemize}

@@

%\input{tex/refactor}

@@

To support the last paragraph, we offer several
specific studies
showing that   widely-held views in SE that are   now questionable, given recent results. 
The first few come from the general SE literature while the rest
relate specifically to bad smells:

\begin{enumerate} 
\item
Nagappan et al. analyzed~\cite{mei15} 
how   {\tt goto}  was used  in 11,000 Github repositories. 
They found that developers
rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
\item
In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em fixing a requirements error  during
coding can be thousands of times more expensive than recognizing and fixing that error
as soon as it arises}. 
We found no evidence for such large   phase delay effect in modern
software (in 171 software projects shepherded
by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ software development languages, tools and development practices. 
Whatever the reason, the main point  is that phase delay is a widely
held belief, despite little recent evidence to support it.
\item
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, They found that  ``(a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
\end{enumerate}
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
careers.

@@

\section{Learning Bad Smell Thresholds}
The SE literature offers  two ways to 
learning bad smell thresholds:
\begin{itemize}
    \item Approaches based on {\em outlier statistics}~\cite{erni96,bender99}
    as used by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
    and Hermans et al.~\cite{hermans15};
    \item  Approaches based on {\em cluster deltas} that we developed
    for   Centroid Deltas~\cite{me12c} and XTREE.
\end{itemize}
\subsection{  Outliers}

The outlier approach assume that unusually large measurements are risk-prone.
Hence, they generate one bad smell threshold for any metric
with such an ``unusually large'' measurements. 
The literature lists several ways to define ``unusually large''.

\subsubsection{Enri \& Lewerentz}
One of the first methods was proposed Erni and Lewerentz~\cite{erni96}. 
Given classes described with the  code metrics of \fig{ck},
Enri and Lewerentz finds the   mean $\mu$ and the standard deviation $\sigma$
of each
code metrics. Their definition of problematic outlier was any code
metric with a measurement greater than $\mu+\sigma$.

Shatnami and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
using $\mu+\sigma$. since it does not   consider the fault-proneness of classes when the thresholds are computed. Also, there is a lack of empirical validation of this method.

\subsubsection{ Shatnami}
Shatnami~\cite{Shatnawi10}'s preferred alternative to $\mu+\sigma$
is to use the VARL method (Value of Acceptable Risk Level) proposed by Bender~\cite{bender99} in his epidemiology studies.  This approach uses two
thresholds which, following Shatnami's guidance, we set to
$p_0=p_1=0.05$. VARL encodes the defect count
for each class to 0 (for no defects known) to 1 (for defects known).
Univariate binary logistic regression is applied to learn three coefficients: $\alpha, \beta,p_0$ being
the  constant intercept ($\alpha$), the coefficient for maximizing log-likelihood function ($\beta$),
and a measure of how well this regression model predicts for the defects ($p_0$).
Any code metric with $p_0>0.05$ is then rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
the equation proposed by Bender:
\begin{equation}
 \mathit{bad\; smell\; if\;} M_c > \frac{1}{\beta }\left( {\log \left( {\frac{{{p_1}}}{{1 - {p_1}}}} \right) - \alpha } \right) 
\end{equation}
\subsubsection{ Alves et al.}
Alves et al.~\cite{Alves2010} propose another approach
to finding thresholds that  utilizes the underlying statistical distribution and scale of the metrics. 
Evey metric value is weighted according to the source lines of code (LOC) of the class. All the weighted metrics are then normalized; i.e. they are divided by the sum of all weights of the same system. As with Shatnawi, univariant regression
is then applied to reject code metrics that are poor predictors of defects (i.e. all those  with $p > 0.05$).
Following this, the normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
Alves et al. then compute their 
thresholds are then derived by choosing the percentage of the overall code that needs to be represented. For instance, Alves et al. suggest the use 90\% quantile of the overall code to derive the threshold for a specific metric. The intuition here is that the worst
code are the outliers beyond 90\% of the normal code measurements.

Our reading of the literature is that the Alves et al.  
approach is preferred to that of Shatnami. For example, Herman et al.~\cite{hermans15} used the
Alves et al. method in their  2015 paper on
exploring bad smells.

\subsubsection{Discussion of the Outlier Approach}
The advantage of the outlier
approach is that it is simple to implement. 
For example, in less than a day we could
code up and apply all the approaches of Shatnami and 
Alves et al.

On the other hand, these outlier approaches have major two disadvantages. 
Firstly, they are {\em verbose}; i.e.  they can
propose thresholds on most of the code metrics.
Such verbosity makes it hard   to critique and prune bad smells
since any bad smell mentioned by a developer has some support
somewhere amongst all those proposed thresholds.

Secondly, the outlier approaches described above are all 
{\em univariant} which means
they do  not understand
that sets of code metrics must
be changed at the same time.
For example:
\begin{itemize} 
\item One way to remove unusually large methods is to reorganize that method's
functionality into several smaller ones. 
\item If that is done,
this means that {\em reducing} lines of code in one method
will {\em increase} the coupling of the methods of that class to
either itself or to other classes (depending on whether or not
the new smaller methods are placed in the original class, or in other classes).
\item
Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
large lines of code {\em and} seemingly excessive coupling since
one cannot be reduced without increasing the other.
\end{itemize}
\subsection{Cluster Deltas}
Borges and Menzies first proposed the centroid delta approach to avoid the drawbacks of univariant reasoning.
Centroid deltas are a {\em multivariant} method which
work  as follows:
\begin{itemize}
    \item Cluster data about code modules;
    \item Find
neighboring clusters $C_+,C_-$ where $C_+$ has more examples of defective
modules than $C_-$;
\item Compute the  delta   in code metrics between the clusters using \mbox{$\Delta = C_+ - C_-$};
\item $\Delta$ are changes needed in defective modules of $C_+$ to
      make them more like the less-defective modules of $C_-$
\end{itemize}
Note that $\Delta$ is a multivariant recommendation
that proposes multiple changes in order to reduce defects.
Since it is computed
from neighboring clusters, it   reasons about
examples with similar distributions.  Hence,   $\Delta$  respects the naturally occurring constraints in the data. For example,
repeating the above example about large methods,   $\Delta$   will not  suggest lowering lines of code
without also increasing a coupling measure.

The drawback with with CD is that it is very {\em verbose}
since
CD   recommended changes to all code
metrics with different values in those two centroids. 
This makes it hard to use CD to   critique and prune away bad smells.
\subsubsection{XTREE}

  XTREE  is cluster delta algorithm
  that avoids the problem of verbose $\Delta$s.
  Instead of reasoning over cluster centroids,
  XTREE utilizes a decision tree learning approach
  to find the fewest differences groupings of examples.

  Decision trees are built via
  {\em iterative dichotomization} (ID).
  ID   finds  a  split  in  the  values  of  independent  features  that  most  reduces the variability of the dependent feature. The data is then divided on that split
  and ID is called
  recursively on the data in each split. In this way,
  each split is the root of a sub-tree.

  The CART iterative dichotomization algorithm
  measures the variability of continuous attribites
  using standard deviation 
  $\sigma_x =\sqrt{ \sum^n_{i=1} \left(  \frac{(x_i - \bar{x})^2}{n-1}\right)$.
  If a code metric is split into $r$ ranges and each range is of
  size
    $n_r$ is
  associated with a set of defect counts  with standard deviation
  $\sigma_r$ then the best split is the one that minimizes
  $\sum_r\frac{n_r}{n}\sigma_x$ (where $n=\sum_r n_r$).

  \begin{figure}[t!!]
\centering
\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
\caption{Sample plans from XTREE. This
tree can be read like a nested if-then-else statement. For
example, lines 3 and 8 show two branches for lines of code
(denoted here as {\em \$loc}) below 698 and above 698.
Any line with a colon ":" character  shows
a leaf of this nesting. For example, if some new code module
is passed down this tree and 
falls to the  line marked in orange,
the colon on that line indicates a 
prediction that this module has a 100\% chance of being defective.
XTREE looks for a nearby branch that has a lower chance of being
defective.
Finding the green branch, XTREE a bad smell threshold
for that module that is the delta between 
the orange branch and the green branch. In this case, that threshold
relates to 
lines of code and comments ({\em lcom}) and the cohesion between 
classes ({\em cam}:   measures similarity of
parameter list to assess the relatedness among the methods of class).
\label{fig:xtree_samp}
\end{figure}

  While not usually labelled as such, ID can
  be viewed as a clustering algorithm that groups together code modules
  with similar defect counts and some shared ranges of code metrics.
  XTREE is a post-processor to ID that reports the delta between 
  nearby groups with different defect counts. \fig{xtree_samp} offers
  a small example of that reflection. In this example:
  \begin{itemize}
      \item Some project has a table of data describing its  classes in the form of \fig{ck}.
      \item After code inspections and running test cases or operational
      tests, each such class is augmented with a defect count.
      \item CART takes that table of data and, using iterative dichotomization,
      it generated the tree of \fig{xtree_samp}.
      \item Once the tree is built, some test case arrives and is passed
      down the tree to a leaf note (see the orange line in \fig{xtree_samp}.
      \item XTREE then looks for a nearby leaf node with a lower defect
      count (see the green line in \fig{xtree_samp}. XTREE bad smell
      thresholds for that method is then the difference between 
      green and orange.
  \end{itemize}

\subsubsection{XTREE: the details}

 {\em Step1:} Using the training data,  XTREE uses iterative dichotomization to
  divide a table of $N$ code modules  into  groups of
size $\alpha=\sqrt{N}$.

{\em Step2:} For each test code module, 
	  find the {\em current } leaf: take each test , run it down to a leaf in the decision tree.  
After that,	  find the {\em desired} leaf:
		\begin{itemize}[leftmargin=3mm]
		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
		\item Identify {\em sibling} leaves; i.e. leaf clusters that can be reached from level $lvl$ that are not same as {\em current }
		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\gamma=0.5$ times the mean score of {\em current}. 
		   If none found, then repeat for $lvl += 1$. Also,
		    return no plan if the new $lvl$ is above the root. 
		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
		\ei
	 Also, find the {\em delta}; i.e. the set difference between  conditions in the decision tree branch to {\em desired} and {\em current}. To find that delta: 
	 \begin{itemize}
	     \item For discrete attributes, delta is the value from {\em desired};
	     \item For  numerics, delta is the numeric difference;
	     \item For numerics  discretized into ranges, delta is a random number selected from the low and high boundaries of the that range.
	     \end{itemize}
	     The bad smell threshold for this module is the delta that would move
	     it from the current to desired lead.

@@

XXXX only reorg due to the issues in the data sets.
we are stuck with that. note hwever that if
projects improved their data collection then
our techniques woudl still apply, unmodifed, to the revised data sets. so here we are ding what we can with the data currently available (and also ensuring we are good to go on future data)

### 2016-03-09 tex/refactor.tex
Updates from ShareLaTeX

\begin{figure*}[!t] 
\scriptsize
\centering
\begin{tabular}{r|c|c|c|c|c}
Mostly from Fowler'99~\cite{fowler99}& & & & & Developer\\
 (with a some additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
  Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
  Combinatorial Explosion~\cite{Kerievsky2005} & & & & \checkmark & \\
  Comments & & & 11 & \checkmark & VL\\
  Conditional Complexity~\cite{Kerievsky2005} & & & 14 & \checkmark & ?\\
  Data Class & \checkmark & & & \checkmark &\\
  Data Clumps &  &  & & \checkmark &\\
  Divergent Change & & & & \checkmark & \\
  Duplicated Code & \checkmark & \checkmark & 1 & \checkmark & VH\\
  Feature Envy & \checkmark & & 8 & \checkmark &\\

>>>>

\begin{figure*}[!t] 
\scriptsize
\centering
\begin{tabular}{r|c|c|c|c|c}
Mostly from Fowler'99~\cite{fowler99}& & & & & Developer\\
 (with a some additions~\cite{Kerievsky2005}) & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
  Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
  Combinatorial Explosion~\cite{Kerievsky2005} & & & & \checkmark & \\
  Comments & & & 11 & \checkmark & VL\\
  Conditional Complexity~\cite{Kerievsky2005} & & & 14 & \checkmark & ?\\
  Data Class & \checkmark & & & \checkmark &\\
  Data Clumps &  &  & & \checkmark &\\
  Divergent Change & & & & \checkmark & \\
  Duplicated Code & \checkmark & \checkmark & 1 & \checkmark & VH\\
  Feature Envy & \checkmark & & 8 & \checkmark &\\
